---
title: "R Lecture 1"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

#設定工作目錄

`getwd()`

`setwd("C:\\Program Files\\R\\working")`


#讀取資料
`read.table()`
.txt, 以空白(“ “)或Tab(“\t”)做分隔。

`read.csv()`
.csv, 以，分隔。

#匯出資料

`write.csv(df, "df.csv", sep=",", col.names=TRUE)`

`write.table(df, "df.txt", quote=FALSE, sep="\t")`

#預覽資料
`head()`, 前6筆 

`tail()`, 後6筆

#summarize 資料
`str()` 列出資料內每個欄位的狀態

`summary()` 列出 5-number summary & mean

`names(), colnames()`得到變數名稱

`class()`  得到型態

`table(), mean(), min(), max(), range(), var(), sd(), cov()`
`quantile(), quantile(變數, c(0.3, 0.7, 0.99))`
`dim()` 給幾個row 幾個column

#套件
`install.packages()` 裝套件

`library()` 匯入套件

#向量化運算
`lapply(資料,fnc)`, `sapply()`

#畫圖
- `plot(V1,V2)`

- `library(lattice)`

`xyplot(V2 ~ V1, data)` 以公式的方式來指定 x 與 y 軸的資料

`xyplot(V2 ~ V1 | 類別變數, data)`

- `library(ggplot2)`

`ggplot(data, aes(x = V1, y = V2)) +`

`geom_point()`
 
 http://www.sthda.com/english/wiki/data-visualization

#缺值

`is.na()` 是否有缺值
 
`data[!complete.cases(data),]` list rows of data that have missing values

`newdata <- na.omit(data)` create new dataset without missing data 

`na.omit()` omit all rows containing missing values.

```{r}
data <- c(12, 4, 6, NA, 45, NA,99)
is.na(data) #是否是NA
which(is.na(data)) #哪個是NA
sum(is.na(data)) #幾個NA
mean(data) #有NA無法跑
mean(data, na.rm = TRUE) #除掉NA算
data[is.na(data)] <- mean(data, na.rm = TRUE) #NA用mean取代
data
data[data == 99] <- NA #99改成NA
data
```


#例子-Car Sale
```{r results = "hide"}
setwd("~/Downloads")
print(load("carsale.Rda"))

```
```{r results = "hide"}
str(mydata)
summary(mydata)
class(mydata)
class(mydata[,1])
sapply(mydata, class)

```
```{r}
table(mydata$maker)

```
```{r}
plot(mydata$long,mydata$lat)
```

```{r}
library(lattice)
xyplot(lat ~ long, mydata)
xyplot(lat ~ long | type, mydata)

```
```{r}
#install.packages("ggplot2")
library(ggplot2)
ggplot(mydata, aes(x = long, y = lat)) +
  geom_point()+
  theme_classic()
```

```{r}
ggplot(mydata, aes(x = long, y = lat)) +
  geom_point() +
  facet_wrap(~ type, ncol = 3)
```

```{r}
df_type = mydata[which(mydata$type != "NA"),] #去掉type有NA的
df = na.omit(mydata) #去掉所有NA  
summary(df$city)
table(df$city)
levels(df$city)
```
最貴的車？
```{r}
max(mydata$price,na.rm=T)
mydata[which.max(mydata$price),]$body

median(c(6000,30000))
mydata[which(mydata$price==max(mydata$price,na.rm=T)),]$price=median(c(6000,30000)) #用中位數取代
mydata[which.max(mydata$price),]$header #哪部車最貴
head(sort(mydata$price,decreasing=T))

```
子集合
```{r}
bmw_df <- subset(df, maker=="bmw")
head(bmw_df)
subset(df, price > 40000, select = c(price, maker,type))
subset(df, price > 40000 & type=='SUV', select = c(price, maker, type))

```


